Do return series have power-law tails with the same index?

### 2017-05-02 papers/FX/draft.tex
Removed a paragraph that I think Thomas would want to remove.
The case when b is not zero is not monotonic.

\subsection{Pareto-distributed returns}\label{sec:pareto_tail}
Since we are interested in the influence of heavy-tailed losses on the preferences of an investor 
we assume the following toy model. 
We consider the case when $X$ has a two-sided Pareto \ds\ given by
\beam\label{eq:pareto}
  F_X(x) = \left\{
  \begin{array}{ll}
    p \left(
    {K \over K - x}
    \right)^\alpha & x \leq 0 \,,\\
    1 - (1 - p) \left(
    {K' \over K' + x}
    \right)^\beta & x > 0\,,
  \end{array}
  \right.
\eeam
where {\red where needed?}
$\alpha, \beta > 1$, $0 < p < 1$ and $K,K' > 0$. Then we have
\beam\label{eq:xxie1.0}
 \wt u(\phi,F_X)\nonumber
  &=&
  \alpha\, K^\alpha\,  p\,
  \int_{-\infty}^0
  u\big( (1 - \phi) \ex^r + \phi \ex^x \big)\,
  {1 + b \over (K - x)^{\alpha + 1}} \,dx\\
    &&+
  \beta \,(K')^\beta\, (1 - p)
  \int_{0}^\infty
  u\big( (1 - \phi) \ex^r + \phi\, \ex^x \big)\,
  {1 + b \1(x <  q) \over (K' + x)^{\beta + 1}} \,dx 
  - b \,u(\delta v) \,F_X(q) \,.\label{eq:pp}
\eeam
\par
We observe the following property which also applies to \eqref{eq:xxie1.0}. 
\ble\label{thrm:I}
 Assume the two-sided Pareto model \eqref{eq:pareto} and that the utility
 \fct\ $u$ is increasing and differentiable. Then $\frac{\pd \wt
   u_{\rm max}}{\pd \alpha} > 0$ and $\frac{\pd \wt u_{\rm max}}{\pd
   K} < 0$.
\ele
The proof is given in Appendix~\ref{sec:thrmI_proof}. 
\par
We conclude that
$\wt u_{\rm max}$ increases with $\alpha$ and decreases with $K$. Therefore there is a curve of
equal preference on the $(\alpha, K)$-plane. Moving along this
curve in the direction of increasing $\alpha$, one expects the values
of $K$ to increase too, i.e., the estimated values of $\alpha$ and $K$
should appear positively dependent. This is indeed the case for some
real return data, e.g. the ``Energy'', ``Consumer Staples'' and
``Information Technology'' sectors of the S\&P 500 index; see 
Figure~\ref{fig:sectors_parameters}.
\par
We illustrate Lemma~\ref{thrm:I} in Figure~\ref{fig:preference_pareto}
for the power-utility \fct\ \eqref{eq:u}.
A particularly interesting special case is when the left and the right
tails have equal parameters, i.e. when $\alpha = \beta$ and $K =
K'$. In this case \eqref{eq:pareto} becomes
\begin{eqnarray*}
  \wt u(\phi, F_X) &=& \alpha p K^\alpha (1 + b)
  \int_{0}^\infty {
    u(C(x)) \left[
      1 - {b \over 1 + b}\1_{\{x \ge q\}}
      \right]
    + u(C(-x))
    \over
    (K + x)^{\alpha + 1}
  } dx
  - b u(\delta v) F_X(q)
\end{eqnarray*}
A detailed analysis of the montonicity of
\[
u(C(x)) \left[
  1 - {b \over 1 + b}\1_{\{x \ge q\}}
  \right]
+ u(C(-x))
\]
which comprises lemma \ref{lemma:III}, allows us to apply lemma
\ref{lemma:I} and conclude the following
\begin{itemize}
\item I am a good guy
\item So are you
\end{itemize}

@@

  -{\hat\phi^{-\xi} \over \xi} \left\{

@@

    \left[{(1 - \hat\phi) \over \hat\phi} e^r + e^x\right]^{-\xi}

@@

    \left[{(1 - \hat\phi) \over \hat\phi} e^r + e^{-x}\right]^{-\xi}

@@

  convenience of writing, let $a = (1 - \hat\phi)e^r/\hat\phi$.

@@

    f(y) &=& \log(a + y) - \log(a y + 1) \\
    g(y) &=& {1 -\xi \over 1+\xi} \log(y) + \log(c) \1_{\{y > q\}}
  Observe
  \begin{eqnarray}
    f(y) - g(y) &=& \log\left(
      {a + y \over a y + 1}
      y^{\xi - 1 \over \xi + 1}
    \right) - \log(c)\1_{\{y > e^q\}} \label{eq:f-g}
  \end{eqnarray}
  and when $y \neq e^q$,
  \begin{eqnarray}
    {\pd f(y) \over \pd y} &=& {1 - a^2 \over (a + y) (a y + 1)}
    \label{eq:fd}\\
    {\pd g(y) \over \pd y} &=& {1 - \xi \over (1 + \xi) y}
    \label{eq:gd}
  \end{eqnarray}
  Furthermore

@@

        So $U'(x)$ and hence $U(x)$ as a whole is not monotonic on $(0, \infty)$.

@@

      \item
        If $ \lim_{y \to e^q} f(y) \le \lim_{y \downarrow e^q} g(y)$,
        there is no solution to $f(y) = g(y)$ on $(e^q,
        \infty)$ either. $U'(x)$ is also monotonic on
        $(e^q, \infty)$. By the calculation as before, it can be seen
        ${\pd U'(x) \over \pd x} < 0$ for positive $x$ that is
        sufficiently close to 0; and
        ${\pd U'(x) \over \pd x} > 0$ for sufficiently large $x$.
        So $U'(x)$ is not monotonic in this case.
      \end{itemize}
    \end{enumerate}
  \item If $a < \min\{1, \xi\}$ and $\xi > 1$, i.e. $a < 1 < \xi$, it
    is clear
    \[
    y_- = {
      a^2 - \xi - \sqrt{(a^2 - 1) (a^2 - \xi^2)}
      \over
      a (\xi - 1)
    } < 0
    \]
    It remains to compare $y_+$ with $1$ to determine whether $f(y) =
    g(y)$ has a solution. Assume $y_+ > 1$. Then
    \begin{eqnarray*}
      a^2 - \xi + \sqrt{(a^2 - 1) (a^2 - \xi^2)} &>& a (\xi - 1) \\
      2 a (\xi - 1) (a + 1) (a - \xi) &>& 0
    \end{eqnarray*}
    This contradicts the assumption $a < \xi$ and $\xi > 1$. Hence
    $y_+ < 1$. So we conclude $(f - g)(y)$ is monotonically increasing
    on both $(1, e^q)$ and $(e^q, \infty)$. Following the same
    analysis as in the case (1.a.i), one can see $U'(\cdot)$ is not
    monotonic on $(0, \infty)$.

@@

    If $(f-g)(y_-) > 0$, $(f-g)(y) = 0$ must have a solution on $(y_-,
    \infty)$. So $U'(\cdot)$ is not monotonic.
  \end{enumerate}
\end{proof}

>>>>

\subsection{Pareto-distributed returns}\label{sec:pareto_tail}
Since we are interested in the influence of heavy-tailed losses on the preferences of an investor 
we assume the following toy model. 
We consider the case when $X$ has a two-sided Pareto \ds\ given by
\beam\label{eq:pareto}
  F_X(x) = \left\{
  \begin{array}{ll}
    p \left(
    {K \over K - x}
    \right)^\alpha & x \leq 0 \,,\\
    1 - (1 - p) \left(
    {K' \over K' + x}
    \right)^\beta & x > 0\,,
  \end{array}
  \right.
\eeam
where {\red where needed?}
$\alpha, \beta > 1$, $0 < p < 1$ and $K,K' > 0$. Then we have
\beam\label{eq:xxie1.0}
 \wt u(\phi,F_X)\nonumber
  &=&
  \alpha\, K^\alpha\,  p\,
  \int_{-\infty}^0
  u\big( (1 - \phi) \ex^r + \phi \ex^x \big)\,
  {1 + b \over (K - x)^{\alpha + 1}} \,dx\\
    &&+
  \beta \,(K')^\beta\, (1 - p)
  \int_{0}^\infty
  u\big( (1 - \phi) \ex^r + \phi\, \ex^x \big)\,
  {1 + b \1(x <  q) \over (K' + x)^{\beta + 1}} \,dx 
  - b \,u(\delta v) \,F_X(q) \,.\label{eq:pp}
\eeam
\par
We observe the following property which also applies to \eqref{eq:xxie1.0}. 
\ble\label{thrm:I}
 Assume the two-sided Pareto model \eqref{eq:pareto} and that the utility
 \fct\ $u$ is increasing and differentiable. Then $\frac{\pd \wt
   u_{\rm max}}{\pd \alpha} > 0$ and $\frac{\pd \wt u_{\rm max}}{\pd
   K} < 0$.
\ele
The proof is given in Appendix~\ref{sec:thrmI_proof}. 
\par
We conclude that
$\wt u_{\rm max}$ increases with $\alpha$ and decreases with $K$. Therefore there is a curve of
equal preference on the $(\alpha, K)$-plane. Moving along this
curve in the direction of increasing $\alpha$, one expects the values
of $K$ to increase too, i.e., the estimated values of $\alpha$ and $K$
should appear positively dependent. This is indeed the case for some
real return data, e.g. the ``Energy'', ``Consumer Staples'' and
``Information Technology'' sectors of the S\&P 500 index; see 
Figure~\ref{fig:sectors_parameters}.
\par
We illustrate Lemma~\ref{thrm:I} in Figure~\ref{fig:preference_pareto}
for the power-utility \fct\ \eqref{eq:u}.
A particularly interesting special case is when the left and the right
tails have equal parameters, i.e. when $\alpha = \beta$ and $K =
K'$. In this case \eqref{eq:pareto} becomes
\begin{eqnarray*}
  \wt u(\phi, F_X) &=& \alpha p K^\alpha (1 + b)
  \int_{0}^\infty {
    u(C(x)) \left[
      1 - {b \over 1 + b}\1_{\{x \ge q\}}
      \right]
    + u(C(-x))
    \over
    (K + x)^{\alpha + 1}
  } dx
  - b u(\delta v) F_X(q)
\end{eqnarray*}
First of all we note that the function
\[
u(C(x)) \left[
  1 - {b \over 1 + b}\1_{\{x \ge q\}}
  \right]
+ u(C(-x))
\]
is not monotonic. To see this we may re-write the above as
\begin{equation}
  -{\phi^{-\xi} \over \xi} \left\{
  \underbrace{
    \left(1 - {b \over 1 + b}\1_{\{x \ge q\}}\right)
    \left[{(1 - \phi) \over \phi} e^r + e^x\right]^{-\xi}
    +
    \left[{(1 - \phi) \over \phi} e^r + e^{-x}\right]^{-\xi}
  }_{U(x)} \right\}
  \label{eq:frg}
\end{equation}
Direct computation gives
\begin{eqnarray*}
  {\pd U(x) \over \pd x}
  &=&
  -{
    (a + e^x)^{-\xi - 1} \xi e^x
    \over
    1 + b \1_{\{x > q\}}
  } + (a + e^{-x})^{-\xi - 1} \xi e^{-x}
  \quad x \neq q
\end{eqnarray*}
where $a = (1 - \phi) e^r/\phi$. For $x \ge q$, we have
\begin{eqnarray*}
  {1 \over \xi}
  {\pd U(x) \over \pd x} &=& {x' \over (a + x')^{\xi + 1}}
  - {x'^\xi \over (a x' + 1)^{\xi + 1} (1 + b)}
\end{eqnarray*}
where we have substituted $x'$ for $e^{-x}$. When $x$ is very large
i.e. when $x'$ is very close to 0, it is useful to expand the last
expression at $x' = 0$. This yields
\[
  {1 \over \xi}
  {\pd U(x) \over \pd x}
  = {x' \over a^{\xi + 1}} + o(x')
\]
So we see that when $x$ is sufficiently large
${\pd U(x) \over \pd x} > 0$. On the other hand, at the discontinuity
point $x = q$, we have
\[
\lim_{x\downarrow q}U(x) - \lim_{x\uparrow q}U(x) = - {
  (a + e^q)^{-\xi} b
  \over
  1+ b
} < 0
\]
So $U(x)$ and hence \eqref{eq:frg} is not monotonic.

which comprises lemma \ref{lemma:III}, allows us to apply lemma
\ref{lemma:I} and conclude the following
\begin{itemize}
\item I am a good guy
\item So are you
\end{itemize}

@@

  -{\phi^{-\xi} \over \xi} \left\{

@@

    \left[{(1 - \phi) \over \phi} e^r + e^x\right]^{-\xi}

@@

    \left[{(1 - \phi) \over \phi} e^r + e^{-x}\right]^{-\xi}

@@

  convenience of writing, let $a = (1 - \phi)e^r/\phi$.

@@

  $f(y)$, $g(y)$ as above. Observe
  % \begin{eqnarray}
  %   f(y) - g(y) &=& \log\left(
  %     {a + y \over a y + 1}
  %     y^{\xi - 1 \over \xi + 1}
  %   \right) - \log(c)\1_{\{y > e^q\}} \label{eq:f-g}
  % \end{eqnarray}
  % and when $y \neq e^q$,
  % \begin{eqnarray}
  %   {\pd f(y) \over \pd y} &=& {1 - a^2 \over (a + y) (a y + 1)}
  %   \label{eq:fd}\\
  %   {\pd g(y) \over \pd y} &=& {1 - \xi \over (1 + \xi) y}
  %   \label{eq:gd}
  % \end{eqnarray}
  % Furthermore

@@

        So $U'(x)$ and hence $U(x)$ as a whole is not monotonic on
        $(0, \infty)$.

@@

      \item
        If $ \lim_{y \to e^q} f(y) \le \lim_{y \downarrow e^q} g(y)$,
        there is no solution to $f(y) = g(y)$ on $(e^q,
        \infty)$ either. $U'(x)$ is also monotonic on
        $(e^q, \infty)$. By the calculation as before, it can be seen
        ${\pd U'(x) \over \pd x} < 0$ for positive $x$ that is
        sufficiently close to 0; and
        ${\pd U'(x) \over \pd x} > 0$ for sufficiently large $x$.
        So $U'(x)$ is not monotonic in this case.
      \end{itemize}
    \end{enumerate}
  \item If $a < \min\{1, \xi\}$ and $\xi > 1$, i.e. $a < 1 < \xi$, it
    is clear
    \[
    y_- = {
      a^2 - \xi - \sqrt{(a^2 - 1) (a^2 - \xi^2)}
      \over
      a (\xi - 1)
    } < 0
    \]
    It remains to compare $y_+$ with $1$ to determine whether
    ${\pd (f -g)(y) \over \pd y} = 0$ has a solution on $(1, \infty)$.
    Assume $y_+ > 1$. Then
    \begin{eqnarray}
      a^2 - \xi + \sqrt{(a^2 - 1) (a^2 - \xi^2)} &>& a (\xi - 1)
      \nonumber \\
      2 a (\xi - 1) (a + 1) (a - \xi) &>& 0 \label{eq:mi6}
    \end{eqnarray}
    This contradicts the assumption $a < \xi$ and $\xi > 1$. Hence
    $y_+ < 1$. So we conclude $(f - g)(y)$ is monotonically increasing
    on both $(1, e^q)$ and $(e^q, \infty)$. Following the same
    analysis as in the case (1.a.i), one can see $U'(\cdot)$ is not
    monotonic on $(0, \infty)$.

@@

    If $(f-g)(y_-) > 0$, $(f-g)(y) = 0$ must have a solution on $(y_-,
    \infty)$. So $U'(\cdot)$ is not monotonic.

    \item $a > \max\{1, \xi\}$ and $\xi > 1$, i.e. $1 < \xi < a$. By
      the same argument that leads to \eqref{eq:mi6}, we see
      $y_+ > 1$. From \eqref{eq:f-g-der} it is clear $(f-g)(y_+)$ is a
      minimum. If $(f-g)(y_+) > 0$, there is no solution to
      $(f-g)(y) = 0$ on $(1, \infty)$; if $(f-g)(y_+) < 0$,
      there must be a solution to $(f-g)(y) = 0$ on $(y_+, \infty)$.

  \end{enumerate}
\end{proof}
